\appendix
\subsection{Appendix A}
\subsubsection{Interview procedure and questions}

\emph{Setting up a user study session}


\begin{enumerate}
     \item The conductor starts a meeting in Zoom.
     \item The conductor explains the steps to the participant.
     \item When consent of recording the session is acquired by the conductor, the conductor starts recording.
     \item The conductor starts the tool and shares the screen.
     \item The conductor allows the participant to control the conductors machine via Zoom remote control.
     \item The conductor instructs the participant to perform the test.
 \end{enumerate}

\paragraph{Instructions}

The task is to design at least two rooms in a dungeon world, with each variant of the AI.

\begin{itemize}
    \item Step 1: Choose the LOW level of the AI and click “Create World”.

\item Step 2: You are now in the World Editing View. Edit as you please. To enter the room editing view, double click the room you wish to edit.

\item Step 3: You are now in the Room Editing View. Use the brushes on the left to edit the room. Click “End Turn” to end your turn, and let the AI contribute. 

\begin{itemize}
    \item AI-placed tiles are tinted purple. AI-suggested tiles are tinted green.
    \item If you are in the LOW variation of AI, click the green suggestions you’d like to place. Click continue when you want to have your turn again.
    \item To go back to the World Editing View, click “Go To World View”.
\end{itemize}

\item Step 4: When you feel satisfied with your creation, tell me so. Restart the program. Start over at step 1 for the next AI version until all three are used once.

\end{itemize}


\paragraph{Interview Questions}

\begin{itemize}
    \item Q1: Which of the three versions of AI did you prefer?
Why?

\item Q2: Which of the three versions of AI did you find least appealing? 
Why?


\item Q3:  How would you describe the creative experience?


\item Q4: What is your perception of the AI’s behaviour? 


\item Q5: Did you feel your creativity was constrained when using any of the three AIs?


\item Q6: Did you adapt to the different AI versions? 
In what ways?


\item Q7: Did you perceive that the AI adapted to you?
In what ways?


\item Q8: How would you describe the relationship between designer and the AI?


\item Q9: How did the AI’s decisions affect your creative process?


\item Q10: How did the different versions affect your design goals?


\item Q11: What do you think is missing or needs to be improved for an AI as the one of the HIGH-version (with high initiative) to be used in collaborative tools?
\end{itemize}

